Clustering based on probability distributions with application on residential customers

### 2021-10-12 main.Rmd
changed simulation part a bit

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
```

@@

<!-- significance of load clustering -->
Large spatio-temporal data sets, both open and administrative, offer up a world of possibilities for energy and social science research. <!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO. The project provides half-hourly data from over 13,000 household electricity smart meters for nearly two years. <!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Electricity utilities can utilize these smart meter usage patterns of customers to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.

@@

@@

@@

An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates.
The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. The algorithm most ubiquitously employed is K-Means. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance. However, it does not explain the cluster composition by combining it with external data. Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf). Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series. As the methods are often situation specific, 
it makes sense to compare them on the performance rather than any standard performance metric. A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion. @Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models. Most time-series clustering models are limited to handling time domain with same start and end date and time. Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series.

@@

To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts. We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods. A continuous measured variable $y$ of length $T$ indexed by ${0, 1, \dots T-1}$. Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively. Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$. These categories could be integers or some more meaningful labels. For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$. We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series. A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series. The method could preform differently with different jumps between consecutive categories. So a difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered. The code for creating these designs can be found in the Supplementary section (link to github repo). The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$. The rest of the results could be found in the supplementary paper.

@@

Each of the categories $g10,g11$, $g20, g21, g22$, and $g30, g31, g32, g33, g34$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs. As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process. Time series data include basic properties such as trend, seasonality, and auto-correlation, and it is reasonable to construct a time series using these features. However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations. Because the time span during which an entity is observed in order to ascertain its behavior is very short, it is assumed that the time series will remain stationary throughout the observation period. The data type is set to "continuous," and the setup is assumed to be gaussian. For example, in a null design, all categories are considered to be normal (0,1). The mean of the subsequent categories are incremented or decremented from this distribution based on the mean difference between them in non-null designs.

@@

**Three signifiant granularities**

@@

Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs. That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change. We consider different distributions across categories (as in Table \ref{tab:tab-distribution}) that will lead to different designs (as in Table \ref{tab:tab-design}). Figure \ref{fig:plot-3gran} shows the linear and cyclic representation of the simulated variable under these five designs. For the consequent data designs, only graphical representations will be provided.

```{r tab-distribution}
granularity = c("g1", "g2", "g3")
alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
                      "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
                      "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")

@@

knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
```

@@

```{r tab-design}
design = c("design-1", "design-2", "design-3", "design-4", "design5")
g1 = c("null", "alternate", "null", "null", "alternate")
g2 = c("null", "null", "alternate", "null", "alternate")
g3 = c("null", "null", "null", "alternate", "alternate")
table <- tibble(design, g1, g2, g3) %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
table

@@

```{r plot-3gran , out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown. In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design. Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}

@@

**Two significant granularities**

@@

**One signifiant granularity**

@@

data_bind

@@

When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity. This happens when the effects of the two granularities on the measured variable are not additive. Consider a case with just two interacting granularities $g1$ and $g3$ of interest. As opposed to the last case, where we could play with the distribution of $(2+5)$ categories, with interaction we can play with the distribution of $(2*5)$ combination of categories.

@@

All the methods were fitted to each data designs and results are reported through confusion matrices. With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs. For difference=1, the performances are pretty bad for js-robust methods and wpd method. The performance starts getting better with increasing difference and get worse with increasing number of replications. Length of series do not show to have any effect on the performance of the methods. It does not depend on if time series is ar or arma.

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=TRUE, messages=FALSE, warning=FALSE)
# Make sure you have the latest version of rmarkdown and bookdown
#devtools::install_github("rstudio/rmarkdown")
#devtools::install_github("rstudio/bookdown")
library(ggplot2)
library(gravitas)
library(gracsr)
library(ggdendro)
library(dplyr)
library(readr)
library(visdat)
library(ggplot2)
library(tidyverse)
library(naniar)
library(here)
library(tsibble)
library(knitr)
library(patchwork)
library(GGally)
library(distributional)
```

@@

<!-- significance of load clustering -->
Large spatio-temporal data sets, both open and administrative, offer up a world of possibilities for energy and social science research. <!--Beyea (2010) has pointed out, there has been little exploration of the full potential of these data bases and their benefits can reach beyond the original intentions for collecting these data.--> One such data sets for Australia is the Smart Grid, Smart City (SGSC) project (2010–2014) available through [Department of the Environment and Energy](https://data.gov.au/data/organization/doee) and Data61 CSIRO. The project provides half-hourly data from over 13,000 household electricity smart meters for nearly two years. <!---The massive amount of data generated in such projects could be overwhelming for analysis.--> Raw data of load time series can quickly become overwhelming and hard to interpret, necessitating high level representation as load profiles via load clustering. Electricity utilities can utilize the smart meter usage patterns to develop targeted tariffs for individual groups and alleviate the problem of volatility in production by capitalizing on the flexibility of consumers.

@@

A lot of work has been done in the literature on clustering customers by their load characeristics. The key to structured aggregation is to summarize big data of large number of customers into manageable subsets of similar load curves. Customers should be clustered uniquely, with behaviors represented with distinctive profile patterns. Approaches could be deterministic or stochastic. Deterministic approaches are hard to generalize due to the need for detailed knowledge of occupancy, appliances, and their time of use. In contrast, stochastic load analysis does not require extensive data
collection other than the time series itself [13,14]. This technique
suggests viewing load as a random process, where extracting time-series features, or a model of the series, facilitates unsupervised clustering. Any standard clustering methodology consists of the following steps: (a) defining distance (dissimilarity) and similarity mostly through feature extraction and selection; and (b) choosing the clustering algorithm design.
Popular distance measures could be time-domain based or frequency-domain (fast Fourier transform). Any model based extraction works by transforming the series into other other objects such as structure or set of parameters which can be more easily characterised and clustered. A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in [42].In contrast to models, a feature-based extraction is used to explicitly define or automatically extract the curves’ key time features, for instance by application of PCA on the daily curves. @Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings. None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data.
The algorithm most ubiquitously employed is K-Means. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf). Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series.

<!--  The time-domain based feature extraction techniques, such as principal component analysis (PCA), detect key features of daily curves and reasonably link them to -->
<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components. Methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27]. -->

<!-- Time-domain analysis is preferred to specify time characteristics, while methods such as fast Fourier transform are used to reveal frequencies that highlight recurrent daily and annual load variations [27]. Popular time-domain techniques, such as principal component analysis -->
<!-- (PCA), detect key features of daily curves and reasonably link them to -->
<!-- distinctive demand behaviours [28,29]. For instance, PCA can help -->
<!-- reduce diurnal dimensions, e.g., 48 half-hourly power measurements, down to a few orthogonal components. @Tureczek2017-pb presented a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings. None of the $34$ selected papers which focus on clustering consumption are based on Australian smart meter data. -->
<!-- The algorithm most ubiquitously employed is K-Means. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf). Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series. -->

<!-- The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. -->

<!-- The key challenge is the application of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters. Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal time-series common in residential datasets.  -->

<!-- The literature -->
<!-- suggests the clustering of residential customers by their load characteristics. The key challenge is the application -->
<!-- of appropriate processes to reduce the extreme dimensionality of load time series to facilitate unique clusters. -->
<!-- Time feature extraction is a potential remedy, however, it is limited by the type of noisy, patchy, and unequal -->
<!-- time-series common in residential datasets. In this paper we propose a strategy to alleviate these limitations by -->
<!-- converting any types of load time series into map models that can be readily clustered. -->

@@

<!-- The basic idea behind the clustering algorithm based on distribution is that the data, generated from the same distribution, belongs to the same cluster if there exists several distributions in the original data.  -->

@@

<!-- An aggregation approach to reduce the data size is not exactly ideal when dealing with highly heterogeneous profiles, as (XXX big data to the rescue example) demonstrates. -->
<!-- The foundation for this study is Tureczek2017-pb, which conducts a systematic review of the current state of the art in smart meter data analytics, which evaluates approximately 2100 peer-reviewed papers and summarizes the main findings. None of the 34 selected papers which focus on clustering consumption are based on Australian smart meter data. The clustering is frequently applied directly to the raw data without scrutinizing for auto correlation and periodicity. The algorithm most ubiquitously employed is K-Means. But the omission of the time series structure or correlation in the analysis while employing K-Means leads to inefficient clusters. Principal Component Analysis or Self-Organizing Maps removes correlation structures and transitions the data to a reduced feature space, but it comes at a cost of interpretability of the final results. @Tureczek2018-ha has shown that a transformation of data to incorporate autocorrelation before K-Means clustering can improve performance and enable K-Means to deliver smaller clusters with less within-cluster variance. However, it does not explain the cluster composition by combining it with external data. Some papers present pre-processing of the smart-meter data before clustering through principal component analysis or factor analysis for dimensionality reduction or self-organizing maps for 2-Dimensional representation of the data (@Ndiaye2011-pf). Other algorithms used in the literature include k-means variations, hierarchical methods and k-medoids based on a greedy algorithm have been designed to select typical periods in the time series. As the methods are often situation specific,  -->
<!-- it makes sense to compare them on the performance rather than any standard performance metric. A type of clustering based on information theory such as Shannon or Renyi entropy and their variants are addressed in , which differs from typical methods adopted for electricity consumer classification, based on the Euclidean distance notion. @Motlagh2019-yj presents strategy to address the problems on patchy, and unequal time-series common in residential data sets by converting load time series into map models. Most time-series clustering models are limited to handling time domain with same start and end date and time. Most of the solutions to handle this like longest common subsequence, dynamic time warping are prone to computational limit with increased length of the series. -->

@@

To validate the clustering approaches, we create data designs that replicate prototype behaviors that might be seen in electricity data contexts. We spiked several features in the data to see where one method works better than the other and where they might give us the same outcome or the effect of missing data <!---and trends--> on the proposed methods. A continuous measured variable $y$ of length $T$ indexed by $\{0, 1, \dots T-1\}$. Three circular granularities $g1$, $g2$ and $g3$ are considered with $2$, $3$ and $5$ levels respectively. Categories of $g1$, $g2$ and $g3$ are represented by ${g10,g11}$, ${g20, g21, g22}$ and ${g30, g31, g32, g33, g34}$. These categories could be integers or some more meaningful labels. For example, the granularity "day-of-week" could be either represented by ${0, 1, 2, \dots, 6}$ or ${Mon, Tue, \dots, Sun}$. We created independent replications $R = \{25, 250, 500\}$ of all data designs to see if our proposed clustering approaches can detect distinct designs in various groups for small, medium and large number of series. A sample size of $T=\{300, 1000, 5000\}$ is used in all designs to test small, medium and large sized series. The method could preform differently with different jumps between consecutive categories. So a mean difference of $diff = \{1, 2, 5\}$ for corresponding categories are also considered. The code for creating these designs can be found in the Supplementary section (link to github repo). The results for $T=300$ and $R=25$ is shown, that means we have $25$ time series each with length $300$. The rest of the results could be found in the supplementary paper.

@@

Each of the $7$ categories from $\{g10,g11$, $g20, g21,\dots g32, g33, g34\}$ might be assumed to come from the same distribution, a subset of them from the same distribution, a subset of them from separate distributions, or all from different distributions, resulting in various data designs. As the methods ignores the linear progression of time, there is little value in adding time dependency in the data generating process. It is reasonable to construct a time series using properties such as trend, seasonality, and auto-correlation. However, when examining distributions across categories of cyclic granularities, these time series features are lost or addressed independently by considering all seasonal fluctuations. Because the time span during which an entity is observed in order to ascertain its behavior is very short, it is assumed that the time series will remain stationary throughout the observation period. The data type is set to "continuous," and the setup is assumed to be gaussian. For example, in a null design, all categories are considered to be normal (0,1). The mean of the subsequent categories are incremented or decremented from this null design based on the mean difference between them in non-null designs.

@@

_All signifiant granularities_

@@

Consider the scenario when all three granularities $g1$, $g2$, and $g3$ are responsible for distinguishing the designs. That means that for at least one among the to-be-grouped designs, the pattern for each of $g1$, $g2$, and $g3$ will change. We consider different distributions across categories (as in Table \ref{tab:tab-dist-design}) (top) that will lead to different designs (as in Table \ref{tab:tab-dist-design}) (below). Figure \ref{fig:plot-3gran-new} shows the linear and cyclic representation of the simulated variable under these five designs. For the consequent data designs, only graphical representations will be provided.

```{r tab-distribution}
granularity = c("g1", "g2", "g3")
# alternate_design =  c("g10 ~ N(0, 1), g11 ~ N(2, 1)",
#                       "g21 ~ N(2, 1), g22 ~ N(1, 1), g23 ~ N(0, 1)",
#                       "g31 ~ N(0, 1), g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1), g35 ~ N(0, 1)")

alternate_design =  c("g11 ~ N(2, 1)",
                      "g21 ~ N(2, 1), g22 ~ N(1, 1)",
                      "g32 ~ N(1, 1), g33 ~ N(2, 1), g34 ~ N(1, 1)")

@@

# knitr::kable(tab_dist, caption = "Alternate distributions of different categories if they deviate from null.")
```

@@

```{r tab-design}
design = c("design-1", "design-2", "design-3", "design-4", "design5")
g1 = c("null", "alternate", "null", "null", "alternate")
g2 = c("null", "null", "alternate", "null", "alternate")
g3 = c("null", "null", "null", "alternate", "alternate")
table <- tibble(design, g1, g2, g3)

# %>% kable(caption = "5 different designs resulting from considering different distributions across categories.")
```

```{r tab-dist-design}

# library(kableExtra)
# 
# knitr::kables(list(kable(caption = "Alternate distributions of different categories if they deviate from null.",
#    tab_dist
#     ) %>% kable_styling(),
#   kable(caption = "5 different designs resulting from considering different distributions across categories.",
#  tibble(design, g1, g2, g3)
#     ) %>% kable_styling()
#   )
# ) %>% kable_styling()
# 
knitr::kable(
  list(tab_dist,  tibble(design, g1, g2, g3)),
  caption = 'Alternate distributions of different categories  (top), 5 designs resulting from different distributions across categories (below)',
  booktabs = TRUE, valign = 't'
)
```

@@

```{r plot-3gran-new, out.width="100%", fig.cap="The linear (left) and cyclic (right) representation of the measured variable is shown. In this scenario, all of $g1$, $g2$ and $g3$ changes across at least one design. Also, it is not possible to comprehend these patterns across cyclic granularities or group similar series just by looking at the linear plots."}

@@

_Few significant granularities_

@@

_One signifiant granularity_

@@

@@

The earlier method could be extended when two granularities of interest interact and we are interested to group subjects based on the interaction of the two granularities. For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year. For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest. As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction, we can play with the distribution of $6$ combination of categories. Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different combination of categories resulting in different designs.

```{r}
## ----designs

sim_null_normal <- function(nxj, nfacetj, mean, sd, w1 = 0, w2=0) {
  rep(distributional::dist_normal(mu = mean, sigma = sd),
      times = nxj * nfacetj
  )
}

sim_varf_normal <- function(nx, nfacet, mean, sd, w1, w2) {
  rep(dist_normal((mean + seq(0, nfacet - 1, by = 1) * w1),
                  (sd + seq(0, nfacet - 1, by = 1) * w2)), each = nx)
}

sim_varx_normal <- function(nx, nfacet, mean, sd, w1, w2) {
  rep(dist_normal((mean + seq(0, nx - 1, by = 1) * w1),
                  (sd + seq(0, nx - 1, by = 1) * w2)), nfacet)
}

sim_varall_normal <- function(nx, nfacet, mean, sd, w1, w2) {
  dist_normal((mean + seq(0, (nx *nfacet - 1), by = 1) * w1),
              (sd + seq(0,  (nx *  nfacet - 1), by = 1) * w2))
}

## ----data

library(hakear)

nx_val = 2 # number of x-axis levels
nfacet_val = 3 # number of facet levels
w1_val = 1 # increment in mean
w2_val = 0 # increment in sd
mean_val = 0 # mean of normal distribution of starting combination
sd_val = 2 # sd of normal distribution of starting combination
quantile_prob_val = seq(0.1, 0.9, 0.1)
ntimes_val = 300

sim_panel_varall <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varall_normal(2, 3, 5, 10, 5, -1.5)
  sim_dist = sim_varall_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
) %>% unnest(data)  %>% 
  rename( "g2" = "id_facet",
          "g1" = "id_x")

sim_panel_varx <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varx_normal(2, 3, 5, 10, 5, -1.5)
  sim_dist = sim_varx_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
) %>% unnest(data)  %>% 
  rename( "g2" = "id_facet",
          "g1" = "id_x")

sim_panel_varf <- sim_panel(
  nx = nx_val, nfacet =  nfacet_val,
  ntimes = ntimes_val,
  # sim_dist = sim_varf_normal(2, 3, 5, 10, 5, 5)
  sim_dist = sim_varf_normal(nx_val, nfacet_val, mean_val, sd_val, w1_val, w2_val)
) %>% unnest(data)  %>% 
  rename( "g2" = "id_facet",
          "g1" = "id_x")

sim_panel_null <- sim_panel(
  nx = nx_val,
  nfacet =  nfacet_val,
  ntimes = ntimes_val,
  sim_dist = distributional
  ::dist_normal(mean_val, sd_val)
) %>% unnest(c(data)) %>% 
  rename( "g2" = "id_facet",
          "g1" = "id_x")

## ----category-plots

p_null <- sim_panel_null %>%
  ggplot(aes(x = as.factor(g2), y = sim_data)) +
  facet_wrap(~g1, labeller = "label_both") +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(a)", round(null, 2))) +
  xlab("g2") +
  theme_bw()+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 0.8, color = "blue") +
  #ggtitle("design1")+
  theme_validation()

p_varf <- sim_panel_varf %>%
  ggplot(aes(x = as.factor(g2), y = sim_data)) +
  facet_wrap(~g1, labeller = "label_both") +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(b)", round(varf, 2))) +
  xlab("g2")+
  theme_bw()+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 0.8, color = "blue") +
  #ggtitle("design2")+
  theme_validation()

p_varx <- sim_panel_varx %>%
  ggplot(aes(x = as.factor(g2), y = sim_data)) +
  facet_wrap(~g1, labeller = "label_both") +
  geom_boxplot()+
  ylab("")  +
  #ggtitle(paste("(c)", round(varx, 2))) +
  xlab("g2")+
  theme_bw()+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 0.8, color = "blue") +
  #ggtitle("design3")+
  theme_validation()

p_varall <- sim_panel_varall %>%
  ggplot(aes(x = as.factor(g2), y = sim_data)) +
  facet_wrap(~g1, labeller = "label_both") +
  geom_boxplot() +
  ylab("") +
  #ggtitle(paste("(d)", round(varall, 2))) +
  xlab("g2")+
  theme_bw()+ stat_summary(
    fun = median,
    geom = 'line',
    aes(group = 1), size = 0.8, color = "blue") +
  #ggtitle("design4") +
  theme_validation()

change_index <- function(data){

index_new <- map(seq_len(ntimes_val), function(i){
  map((seq_len(nx_val*nfacet_val)), function(j)
    {
value = i + (j-1)*ntimes_val
})
  }) %>% unlist()

data_new = data %>%
  ungroup() %>%
  mutate(index_old = row_number(),
         index_new = index_new)

y = data_new[match(index_new, data_new$index_old),]

y <- y %>%
  mutate(time = row_number())

return(y)
}

endbreaks<- nrow(sim_panel_null)

p1 <- change_index(sim_panel_null) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")
  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
    #    panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p2 <- change_index(sim_panel_varf) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
     #   panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p3 <- change_index(sim_panel_varx) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue")+
  #theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
   #     panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p4 <- change_index(sim_panel_varall) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  #geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)

<!-- When two granularities of interest interact, the connection between a granularity and the measured variable is determined by the value of the other interacting granularity. This happens when the effects of the two granularities on the measured variable are not additive. For simplicity, consider a case with just two interacting granularities $g1$ and $g2$ of interest. As opposed to the last case, where we could play with the distribution of $5$ individual categories, with interaction we can play with the distribution of $6$ combination of categories. Consider $4$ designs in Figure \ref{fig:} where different distributions are assumed for different designs to get some distinction across designs. For example, in application, think about the scenario when customers need to grouped basis their joint behavior across hour-of-day and month-of-year. -->

@@

All the methods were fitted to each data designs and results are reported through confusion matrices. With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs. For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT. Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example. It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month. The performance is promising except when the number of observations for a customer is really small. For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.

```{r}
# code in append_3gran_change.R
table <- read_rds("data/append_3gran_change.rds")
table %>% kable()
```

<!-- starts getting better with increasing difference and get worse with increasing number of replications. Length of series do not show to have any effect on the performance of the methods. It does not depend on if time series is ar or arma. -->
